Low Infant Birth Weight in Brazil

### 2024-04-30 code/eda-code/eda-v1.qmd
Project Part 6
Updated tables per peer reviews

```{r}
# Create histogram plot
age_plot <- ML_summary %>% filter(!is.na(Age)) %>% # Filter out NAs
  ggplot(aes(x = Age)) +
  geom_histogram(binwidth = 0.5, width= 0.8, fill="#69b3a2") +
  geom_vline(xintercept = c(15, 49), linetype = "dashed", color = "red") +  # Add vertical lines that mark reproductive age
  ggtitle("Maternal Age") +
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(age_plot)

@@

```{r}
# Create histogram plot
weight_plot <- ML_summary %>% filter(!is.na(Weightgrams)) %>% #filter out NAs
  ggplot(aes(x = Weightgrams)) +
  geom_histogram(binwidth = 25, fill="#69b3a2") +
  geom_vline(xintercept = c(1500, 4000), linetype = "dashed", color = "red") +  # Add vertical lines that mark normal birthweight
  ggtitle("Infant birthweight") +
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(weight_plot)

# Save plot to file
figure_file = here("results","figures", "weight_distribution.png")
ggsave(filename = figure_file, plot=weight_plot, bg="white") #background is white
```
Now infant birth length.
```{r}
# Create histogram plot
length_plot <- ML_summary %>% filter(!is.na(Lengthcentimeters)) %>% #filter out NAs
  ggplot(aes(x = Lengthcentimeters)) +
  geom_histogram(binwidth = 1, fill="#69b3a2") +
  geom_vline(xintercept = c(49), linetype = "dashed", color = "red") +  # Add vertical lines that mark average birth length for both sexes
  ggtitle("Infant birth length") +
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(length_plot)

>>>>

```{r}
# Create histogram plot
age_plot <- ML_summary %>% filter(!is.na(Age)) %>% # Filter out NAs
  ggplot(aes(x = Age)) +
  geom_histogram(binwidth = 0.5, width= 0.8, fill="#69b3a2") +
  geom_vline(xintercept = c(15, 49), linetype = "dashed", color = "red") +  # Add vertical lines that mark reproductive age
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(age_plot)

@@

```{r}
# Create histogram plot
weight_plot <- ML_summary %>% filter(!is.na(Weightgrams)) %>% #filter out NAs
  ggplot(aes(x = Weightgrams)) +
  geom_histogram(binwidth = 25, fill="#69b3a2") +
  geom_vline(xintercept = c(1500, 4000), linetype = "dashed", color = "red") +  # Add vertical lines that mark normal birthweight
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(weight_plot)

# Save plot to file
figure_file = here("results","figures", "weight_distribution.png")
ggsave(filename = figure_file, plot=weight_plot, bg="white") #background is white
```
Now infant birth length.
```{r}
# Create histogram plot
length_plot <- ML_summary %>% filter(!is.na(Lengthcentimeters)) %>% #filter out NAs
  ggplot(aes(x = Lengthcentimeters)) +
  geom_histogram(binwidth = 1, fill="#69b3a2") +
  geom_vline(xintercept = c(49), linetype = "dashed", color = "red") +  # Add vertical lines that mark average birth length for both sexes
  theme_ipsum() +
  theme(plot.title = element_text(size=15))
plot(length_plot)

### 2024-04-30 products/manuscript/Manuscript_Cassia.qmd
Project Part 6
Updated tables per peer reviews

This cross-sectional study uses linear regression models to understand disparities in infant birth weight in relation to maternal skin color or ancestry, age, gestational status (parity or gravidity), and nationality in Rio de Janeiro, Brazil's first public maternity hospital, the Maternidade Laranjeiras (now Maternidade Escola) in the 1920s. I hypothesize that infants born to women of color (either Black or "*preta*" or mixed-race or "*parda*") will have lower birth weights than infants born to White women, whether Brazilian or immigrant, given the historical legacies of slavery, only abolished in 1888, on population health in Brazil.

@@

The journal published obstetricians’ and gynecologists’ clinical observations, analyses of new surgical techniques, and ANM proceedings. Between June 1922 and May 1926, RGO also published the monthly clinical reports of all women treated at the Maternidade Laranjeiras. I was unable to locate vol. 18, nos. 4, 5, 6, 8 (1924) and vol. 20, no. 4 (1926). From the available issues, I recorded the following information, when available, for all patients: patient number, gravidity and parity, skin color, age, nationality, type of delivery (natural, interventionist, operative), maternal outcome (death, discharge, transferal to separate hospital), birth outcome (spontaneous abortion, stillbirth, live birth, or neonatal death), and the mother’s reproductive history. Gravidity refers to the total number of pregnancies a woman has, regardless of duration. Parity refers to a woman’s number of past pregnancies that reached viability and have been delivered, regardless of the number of children [@posner2013]. Please see the `Supplemental_Materials` document for more information on the dataset.

@@

```{r}
#| label: tbl-table2
#| tbl-cap: "Model 1 Results"
#| echo: FALSE
table2 <- readRDS(here::here("results", "tables", "table2_final.rds"))
flex_table2 <- as_flex_table(table2)
flex_table2
```

@@

```{r}
#| label: tbl-table3
#| tbl-cap: "Model 2 Results"
#| echo: FALSE
table3 <- readRDS(here::here("results", "tables", "table3_final.rds"))
flex_table3 <- as_flex_table(table3)
flex_table3
```

@@

```{r}
#| label: tbl-table4
#| tbl-cap: "Model 3 Results"
#| echo: FALSE
table4 <- readRDS(here::here("results", "tables", "table4_final.rds"))
flex_table4 <- as_flex_table(table4)
flex_table4
```

@@

```{r}
#| label: tbl-table5
#| tbl-cap: "Model 4 Results"
#| echo: FALSE
table5 <- readRDS(here::here("results", "tables", "table5_final.rds"))
flex_table5 <- as_flex_table(table5)
flex_table5
```

>>>>

This cross-sectional study uses linear regression models to understand disparities in infant birth weight in relation to maternal skin color (ancestry), age, gestational status (parity), and nationality in Rio de Janeiro, Brazil's first public maternity hospital, the Maternidade Laranjeiras (now Maternidade Escola) in the 1920s. I hypothesize that infants born to women of color (either Black or "*preta*" or mixed-race or "*parda*") will have lower birth weights than infants born to White women, whether Brazilian or immigrant, given the historical legacies of slavery, only abolished in 1888, on population health in Brazil.

@@

The journal published obstetricians’ and gynecologists’ clinical observations, analyses of new surgical techniques, and ANM proceedings. Between June 1922 and May 1926, RGO also published the monthly clinical reports of all women treated at the Maternidade Laranjeiras. I was unable to locate vol. 18, nos. 4, 5, 6, 8 (1924) and vol. 20, no. 4 (1926). From the available issues, I recorded the following information, when available, for all patients: patient number, gravidity and parity, skin color, age, nationality, type of delivery (natural, interventionist, operative), maternal outcome (death, discharge, transferal to separate hospital), birth outcome (spontaneous abortion, stillbirth, live birth, or neonatal death), and the mother’s reproductive history. Gravidity refers to the total number of pregnancies a woman has, regardless of duration. Parity refers to a woman’s number of past pregnancies that reached viability and have been delivered, regardless of the number of children [@posner2013]. Please see the `Supplemental_Materials` document for more information on the dataset. For full citations on the journals from which these data came, see my other publications [@roth2020, @roth2020a].

@@

```{r}
#| label: tbl-table2
#| tbl-cap: "Model 1 Results"
#| echo: FALSE
table2 <- readRDS(here::here("results", "tables", "table2_final.rds"))
flex_table2 <- as_flex_table(table2)
flex_table2 <- flex_table2 %>%
  flextable::set_table_properties(width = .5, align = "center")
flex_table2
```

@@

```{r}
#| label: tbl-table3
#| tbl-cap: "Model 2 Results"
#| echo: FALSE
i = 16.5
w = i*0.3937
font_size <- 10
table3 <- readRDS(here::here("results", "tables", "table3_final.rds"))
flex_table3 <- as_flex_table(table3)
flex_table3 <- autofit(flex_table3)
auto_widths <- dim(flex_table3)$widths/sum(dim(flex_table3)$widths) # calculate column widths
flex_table3 <- width(flex_table3, width=w*auto_widths) # adjust column widths
flex_table3 <- fontsize(flex_table3, size=font_size) # set font size
flex_table3
```

@@

```{r}
#| label: tbl-table4
#| tbl-cap: "Model 3 Results"
#| echo: FALSE
table4 <- readRDS(here::here("results", "tables", "table4_final.rds"))
flex_table4 <- as_flex_table(table4)
flex_table4 <- flex_table4 %>%
  flextable::set_table_properties(width = .5, align = "center")
flex_table4
```

@@

```{r}
#| label: tbl-table5
#| tbl-cap: "Model 4 Results"
#| echo: FALSE
i = 16.5
w = i*0.3937
font_size <- 10
table5 <- readRDS(here::here("results", "tables", "table5_final.rds"))
flex_table5 <- as_flex_table(table5)
flex_table5 <- autofit(flex_table5)
auto_widths <- dim(flex_table5)$widths/sum(dim(flex_table5)$widths) # calculate column widths
flex_table5 <- width(flex_table5, width=w*auto_widths) # adjust column widths
flex_table5 <- fontsize(flex_table5, size=font_size) # set font size
flex_table5
```
